---
title: "COVID Cases over Time"
author: "Ahmed Elsaeyed"
date: "3/10/2022"
output:
  html_document:
    df_print: paged
    fig_caption: true
    css: ae_theme.css
    theme: cosmo

---

1.
## Setting Up Libraries
I will mainly be using tidyr and dplyr here to do the analysis.

```{r setup, include=FALSE}

knitr::opts_chunk$set(class.source = "codechunk")
library(tidyr)
library(dplyr) 
library(RCurl)
library(ggplot2)

```

2.
## Analysis Questions
Thanks to *Moiya Josephs* for the analysis questions:   
1- Find the state with the highest and lowest deaths.  
2- Compare the death rate both before and after the vaccine was released.   
       
     
3.   
## Pull in the Data from Github   
To answer the questions above I will be doing a lot of sorting and group_by + filtering.  

```{r github }
my_git_url <- getURL("https://raw.githubusercontent.com/aelsaeyed/Data607/main/Project2/COVID/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv")
covid_cases_raw <- read.csv(text = my_git_url)
head(covid_cases_raw, 10)
```

4.
## Fixing up and Subsetting Columns
Some of the columns I need to be numerical are in character format, so I will take care of that first. I will also separate the submission_date column into 3 columns, one each for month, day, and year. The commas in the numerical columns had to be removed first.   

```{r cleanup}
cases_cleaned = covid_cases_raw %>% 
  separate(ï..submission_date, into = c("month", "day", "year"), sep="/") %>%  #separate the date values
  mutate(`new_death`= gsub(",", "", new_death) ) %>% #remove commas
  mutate(`new_death`= as.integer(new_death) ) %>%    #cast to int
  mutate(`tot_death`= gsub(",", "", tot_death) ) %>% #remove commas
  mutate(`tot_death`= as.integer(tot_death) ) %>%    #cast to int
  mutate(`year`= as.integer(year)) %>%    #cast the date values as int as well
  mutate(`month`= as.integer(month)) %>% 
  mutate(`day`= as.integer(day)) %>% 
  arrange(desc(year), desc(month), desc(day)) %>% #arrange by dates descending
  select(3,1,2,4,10,13) #Select only the columns I need
  
head(cases_cleaned, 10)
```


















