---
title: "COVID Cases over Time"
author: "Ahmed Elsaeyed"
date: "3/10/2022"
output:
  html_document:
    df_print: paged
    fig_caption: true
    css: ae_theme.css
    theme: cerulean

---

1.
## Setting Up Libraries
I will mainly be using tidyr and dplyr here to do the analysis.

```{css, echo=FALSE}

```

```{r setup, include=FALSE}

knitr::opts_chunk$set(class.source = "codechunk")
library(tidyr)
library(dplyr) 
library(RCurl)
#library(stringr)
```

2.
## Analysis Questions
Thanks to *Moiya Josephs* for the analysis questions:   
1- "Find the state with the highest and lowest deaths."  
2- "Compare the death rate both before and after the vaccine was released."   
       
     
3.   
## Pull in the Data from Github   
For the first question, I will have to use tidyr to separate the "year" column, as it is includes the start year and end year of each show separated by a dash. I will assume the question asks for the highest rated TV show per release year. For the second question, I will assume that "category" refers to the genre column, which will also require separation as each cell of that column contains several genres that the show might fall into. 

```{r github }
my_git_url <- getURL("https://raw.githubusercontent.com/aelsaeyed/Data607/main/Project2/Heros/Dataset_Superhero-TV-Shows.csv")
superhero_tv_shows_raw <- read.csv(text = my_git_url)
head(superhero_tv_shows_raw, 10)
```


























