---
title: "AhmedElsaeyed-607Lab1"
author: "Ahmed Elsaeyed"
date: "2/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Pulling in the data from my github

```{r}
library(RCurl)
my_git_url <- getURL("https://raw.githubusercontent.com/aelsaeyed/Data607/main/vote_predictions.csv")
trump_vote_predictions <- read.csv(text = my_git_url, quote = "")

my_git_url2 <- getURL("https://raw.githubusercontent.com/aelsaeyed/Data607/main/averages.csv")
vote_averages <- read.csv(text = my_git_url, quote = "")
```

### Data Exploration

This dataset contains data pertaining to members of different chambers of congress and how they voted- specifically how many times they voted in line with Trump's position. Alongside this is a score of how they were predicted to have voted. The website shows a little graphic as a column derived using both sets that compares the number of times the Congressperson was predicted to vote with Trump against how many times they actually did. 
```{r}
library(dplyr)
library(tidyr)
voted_summ = summary(trump_vote_predictions)
voted_summ

avg_summ = summary(vote_averages)
avg_summ

```

### Subsetting
Ill be looking only at the averages file, which contains the data describing how the votes actually played out alongside the prediction. I will be cleaning up the "party" column so that it is readable, and filtering for only the latest 2 congresses, and within that congress I only want data from the House of Representatives. 
```{r}
avg_subset = subset(vote_averages, vote_averages$congress == (117 || 116) & vote_averages$chamber == 'house')

avg_subset$party[avg_subset$party == 'R'] <- "Republican"
avg_subset$party[avg_subset$party == 'D'] <- "Democrat"

glimpse(avg_subset)

```

### Subsetting Pt. 2 - Narrowing Down Districts
I narrowed the data down some more to a list of the districts that correspond with New York State. 
```{r}
NY_subset = subset(avg_subset, avg_subset$state == 'NY')
glimpse(NY_subset)
```

### Further Exploration
Some exploration can be done with this data to see how much Trump polarized the voting from congress to congress. To do that, I would try to see the percentage of people who opposed the expectation to vote in line with Trump on policies, and how that percentage changed for the same people between congresses. 



