---
title: "R and SQL"
author: "Ahmed Elsaeyed"
date: "2/13/2022"
output:
  html_document:
    includes:
      in_header: header.html
    css: ae_theme.css
    highlight: pygments
    theme: cerulean
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RPostgreSQL)
library(TTR)
library(dplyr)
```

# Database Connection and Data Retrieval
```{r}
#Create DB Connection using Postgres
dbname <- "movies"
dbuser <- "postgres"
dbpass <- ""
dbhost <- "localhost"
dbport <- 5432
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, host=dbhost, port=dbport, dbname=dbname,user=dbuser, password=dbpass)

#Send queries and create dataframes

#Get the ratings table- a table that simply describes the relationship
#between people and movies.
rquery <- dbSendQuery(
            con, 
            query <- "select * from ratings")
ratings <- fetch(rquery, n=-1)

#Get the movie table
mquery<- dbSendQuery(
            con, 
            query <- "select * from movies") 
movies<- fetch(mquery, n=-1)


#Get the friends table
fquery<- dbSendQuery(
            con, 
            query <- "select * from friends") 
friends<- fetch(fquery, n=-1)
```

# Close Connection
```{r}
#Close Connections
dbDisconnect(con) 
dbUnloadDriver(drv)
```

# Display the Retrieved Data
```{r}
head(ratings, 36)

```

```{r}
head(movies)

```

```{r}
head(friends)

```

# Joining the Tables to Stitch Together Information
I will first join the movies and ratings table to get a table that exposes the movie names. I will then join on the friends table to expose the people's names. Now I have all 36 ratings- one for each friend for each movie. 
```{r joining-for-data-consolidation}
movies_and_ratings = left_join(movies, ratings, by=c("movie_id" = "movie" ) ) 
          
frnds_movies_ratings = left_join(movies_and_ratings, friends, by=c("rater" = "person_id"))
head(frnds_movies_ratings)
```

# Subsetting
I filtered out the records in which the person has no rating for the movie by just including only the ratings >0. 
```{r filtering}
ratings<- frnds_movies_ratings %>% 
  filter(score > 0) %>%
  select(first_name, movie_name, score)
head(ratings)
```

# Reshaping
I wanted a simple table showed the person's name and their rating for each movie in a row. I decided to reshape the table using a library I looked up called reshape2, to basically "pivot" the above table so that the first become the row names, and the movie names are the columns. 
```{r pivoting}
library(reshape2)
reshaped_df = dcast(ratings, first_name ~ movie_name)
head(reshaped_df)
```
This is how I intuitively collected the data, but it made more sense to create several tables like I did above to have two tables representing people and movies, and then one table to represent the relationship between them- the ratings. That way, the "people" and "movie" information can be expanded easily, for example adding more details like gender, address, movie runtime, and NCAA rating, independently. Having made sure that the tables can be independently changed, I can easily re-run the above to include the people's age in my final table (imagine it wasn't originally in the people table)

# Reshaping with New Data
```{r}
ratings_age<- frnds_movies_ratings %>% 
  filter(score > 0) %>%
  select(first_name, movie_name, score, age)
head(ratings_age)
```

```{r pivoting-again}
library(reshape2)
reshaped_df_ages = dcast(ratings_age, first_name + age ~ movie_name, value.var = "score")
head(reshaped_df_ages)
```
